A Bayesian networks approach to infer social changes in burials in northeastern Taiwan during the European colonization period

### 2021-05-01 analysis/paper/paper.Rmd
edit the introduction based on the R1's comments

<!--archaeological specific case study and our model-->
We use a Bayesian ERGM to study social changes in a pericolonial context at Kiwulan, an Iron Age site in northeastern Taiwan (Figure \@ref(fig:figure-KWL-map)), which was occupied from the 14th to 19th centuries. This occupation period includes the time before the European arrival, the presence of the Spanish and the Dutch in the 17th century, and finally a large wave of Han Chinese in the 19th century [@Chen2007]. Our anthropological model proposes that the influence of a colonial power, combined with high local values attached to imported goods, led to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994]. Increased social inequality associated with the use of foreign prestige goods is recognized as a ‘network’ strategy for gaining power [@Blanton1996; @Feinman2000]. Network-based societies are identified in some places in Asia with involvement in long-distance trade [@Carter2015; @Ueda2016; @Liu2006]. The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by Europeans.

>>>>

<!--archaeological specific case study and our model-->
We use a Bayesian ERGM to study social changes in a pericolonial context at Kiwulan, an Iron Age site in northeastern Taiwan (Figure \@ref(fig:figure-KWL-map)), which was occupied from the 14th to 19th centuries. This occupation period includes the time before the European arrival, the presence of the Spanish and the Dutch in the 17th century, and finally a large wave of Han Chinese in the 19th century [@Chen2007]. Our anthropological model proposes that the influence of an indirect colonial power, combined with high local values attached to imported goods, led to increased social inequality due to competition among individuals [@Brumfiel1994; @Clark1994]. Before the arrival of Europeans, some imported goods already entered Kiwulan because of the regional trade network with Chinese traders [@Chen2004]. We assume that the use of imported goods and their values were increased and stressed after the European arrival. Different from Heping dao and Tamsui in northern Taiwan where European forts located, the remote geographical setting of northeastern Taiwan made it less interfered by European colonial power but experienced frequent trade activities stimulated by the Europeans. This paper focuses on the effect of European colonization in a pericolonial context that is not well understood compared to places that were colonial centers [@Acabado2017; @Trabert2017].

The involvement in long-distance exchange is often associated with changes in political-economic strategies from a corporate mode to a network mode, where ambitious individuals are able to build their local power base and personal influence through the acquisition and distribution of foreign prestige goods [@Blanton1996; @Feinman2000; @Klehm2017]. Network-based societies are identified in some places in Asia that were invovled in long-distance trade [@Carter2015; @Ueda2016; @Liu2006]. The observed uneven distribution of burial goods at Kiwulan may be explained as a result of unequal access to trade goods when Indigenous societies became more involved in the complex trade network stimulated by Europeans. Here we define trade goods as prestige goods based on their archaeological contexts where very few of trade goods were found in middens and more in burials, along with the symbol of personal influence described in the Spanish records [@LiandWu2006]. In addition, the spatial pattern of trade goods in residential presents a more concentration pattern statistically after the European presence. The evidence from different archaeological contexts indicates the possible use as prestige goods.
